---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  dpi=200,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```



# COVID19AR

A package for analysing COVID-19 Argentina's outbreak

 <!-- . -->




# Package

| Release | Usage | Development |
|:--------|:------|:------------|
| | [![minimal R version](https://img.shields.io/badge/R%3E%3D-3.6.0-blue.svg)](https://cran.r-project.org/) | [![Travis](https://travis-ci.org/rOpenStats/COVID19AR.svg?branch=master)](https://travis-ci.org/rOpenStats/COVID19AR) |
| [![CRAN](http://www.r-pkg.org/badges/version/COVID19AR)](https://cran.r-project.org/package=COVID19AR) | | [![codecov](https://codecov.io/gh/rOpenStats/COVID19AR/branch/master/graph/badge.svg)](https://codecov.io/gh/rOpenStats/COVID19AR) |
|||[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)|




# How to get started (Development version)

Install the R package using the following commands on the R console:

```R
# install.packages("devtools")
devtools::install_github("rOpenStats/COVID19AR")
```

# How to use it

First add variable with your prefered data dir configuration in `~/.Renviron`. You will receive a message if you didn't. 

```.Renviron
COVID19AR_data_dir = "~/.R/COVID19AR"
```

```{r, library}
library(COVID19AR)
library(ggplot2)
```


# COVID19AR datos abiertos del Ministerio de Salud de la Nación

opendata From Ministerio de Salud de la Nación Argentina

```{r, covid19curator}
log.dir <- file.path(getEnv("data_dir"), "logs")
dir.create(log.dir, recursive = TRUE, showWarnings = FALSE)
log.file <- file.path(log.dir, "covid19ar.log")
lgr::get_logger("root")$add_appender(AppenderFile$new(log.file))
lgr::threshold("info", lgr::get_logger("root"))
lgr::threshold("info", lgr::get_logger("COVID19ARCurator"))

# Data from
# http://datos.salud.gob.ar/dataset/covid-19-casos-registrados-en-la-republica-argentina
covid19.curator <- COVID19ARCurator$new()

dummy <- covid19.curator$loadData()
dummy <- covid19.curator$curateData()
# Dates of current processed file
max(covid19.curator$data$fecha_apertura, na.rm = TRUE)
# Inicio de síntomas

max(covid19.curator$data$fecha_inicio_sintomas,  na.rm = TRUE)

# Ultima muerte
max(covid19.curator$data$fecha_fallecimiento,  na.rm = TRUE)

report.date <- max(covid19.curator$data$fecha_inicio_sintomas,  na.rm = TRUE)

covid19.ar.provincia.summary <- covid19.curator$makeSummary(group.vars = c("residencia_provincia_nombre"))
covid19.ar.provincia.summary.100.confirmed <- covid19.ar.provincia.summary %>% filter(confirmados >= 100)
# Provinces with > 100 confirmed cases
covid19.ar.provincia.summary.100.confirmed$residencia_provincia_nombre
```

```{r, residencia_provincia_nombre}
covid19.ar.summary <- covid19.curator$makeSummary(group.vars = c("residencia_provincia_nombre"))
nrow(covid19.ar.summary)
porc.cols <- names(covid19.ar.summary)[grep("porc", names(covid19.ar.summary))]
kable((covid19.ar.summary %>% filter(confirmados > 0) %>% arrange(desc(confirmados))) %>% 
        select_at(c("residencia_provincia_nombre", "confirmados", "tests", "fallecidos", "dias.fallecimiento",porc.cols)))
```

```{r, sepi_apertura}
covid19.ar.summary <- covid19.curator$makeSummary(group.vars = c("sepi_apertura"))
nrow(covid19.ar.summary)
porc.cols <- names(covid19.ar.summary)[grep("porc", names(covid19.ar.summary))]
kable(covid19.ar.summary %>% filter(confirmados > 0) %>% arrange(sepi_apertura, desc(confirmados)) %>% select_at(c("sepi_apertura", "sepi_apertura", "confirmados", "tests", "internados", "fallecidos",  porc.cols)))

```

```
```{r, residencia_provincia_nombre-sepi_apertura}
covid19.ar.summary <- covid19.curator$makeSummary(group.vars = c("residencia_provincia_nombre", "sepi_apertura"))
nrow(covid19.ar.summary)
porc.cols <- names(covid19.ar.summary)[grep("porc", names(covid19.ar.summary))]

```

```{r, residencia_provincia_nombre-sepi_apertura-plot-tests}
sepi.fechas <- covid19.curator$data %>% 
  group_by(sepi_apertura) %>% 
  summarize(ultima_fecha_sepi = max(fecha_apertura), .groups = "keep")


data2plot <- covid19.ar.summary %>%
                filter(residencia_provincia_nombre %in% covid19.ar.provincia.summary.100.confirmed$residencia_provincia_nombre) %>%
                filter(confirmados > 0 ) %>%
                filter(positividad.porc <=0.6 | confirmados >= 20)

                
data2plot %<>% inner_join(sepi.fechas, by = "sepi_apertura")
dates <- sort(unique(data2plot$ultima_fecha_sepi))

covplot <- data2plot %>%
 ggplot(aes(x = ultima_fecha_sepi, y = confirmados, color = "confirmados")) +
 geom_line() +
 facet_wrap(~residencia_provincia_nombre, ncol = 2, scales = "free_y") +
 labs(title = "Evolución de casos confirmados y tests\n en provincias > 100 confirmados")
covplot <- covplot +
 geom_line(aes(x = ultima_fecha_sepi, y = tests, color = "tests")) +
 facet_wrap(~residencia_provincia_nombre, ncol = 2, scales = "free_y")
covplot <- setupTheme(covplot, report.date = report.date, x.values = dates, x.type = "dates",
                     total.colors = 2,
                     data.provider.abv = "@msalnacion", base.size = 6)
covplot <- covplot + scale_y_log10()
covplot

```

```{r, residencia_provincia_nombre-sepi_apertura-plot-positividad}

covplot <- data2plot %>%
 ggplot(aes(x = ultima_fecha_sepi, y = positividad.porc, color = "positividad.porc")) +
 geom_line() +
 facet_wrap(~residencia_provincia_nombre, ncol = 2, scales = "free_y") +
 labs(title = "Porcentajes de positividad, uso de UCI, respirador y letalidad\n en provincias > 100 confirmados")
covplot <- covplot +
 geom_line(aes(x = ultima_fecha_sepi, y = cuidado.intensivo.porc, color = "cuidado.intensivo.porc")) +
 facet_wrap(~residencia_provincia_nombre, ncol = 2, scales = "free_y")
covplot <- covplot  +
 geom_line(aes(x = ultima_fecha_sepi, y = respirador.porc, color = "respirador.porc"))+
 facet_wrap(~residencia_provincia_nombre, ncol = 2, scales = "free_y")
covplot <- covplot +
 geom_line(aes(x = ultima_fecha_sepi, y = letalidad.min.porc, color = "letalidad.min.porc")) +
 facet_wrap(~residencia_provincia_nombre, ncol = 2, scales = "free_y")

covplot <- setupTheme(covplot, report.date = report.date, x.values = dates, x.type = "dates",
                     total.colors = 4,
                     data.provider.abv = "@msalnacion", base.size = 6)
covplot
```


```{r, residencia_provincia_nombre-sexo}

covid19.ar.summary <- covid19.curator$makeSummary(group.vars = c("residencia_provincia_nombre", "sexo"))
nrow(covid19.ar.summary)
porc.cols <- names(covid19.ar.summary)[grep("porc", names(covid19.ar.summary))]
kable((covid19.ar.summary %>% filter(confirmados >= 10) %>% arrange(desc(confirmados))) %>% select_at(c("residencia_provincia_nombre", "sexo", "confirmados", "internados", "fallecidos",  porc.cols)))

```

```{r, residencia_provincia_nombre-edad.rango}

covid19.ar.summary <- covid19.curator$makeSummary(group.vars = c("residencia_provincia_nombre", "edad.rango"))

 # Share per province
 provinces.deaths <-covid19.ar.summary %>%
    group_by(residencia_provincia_nombre) %>%
    summarise(fallecidos.total.provincia = sum(fallecidos), .groups = "keep")
 covid19.ar.summary %<>% inner_join(provinces.deaths, by = "residencia_provincia_nombre")
 covid19.ar.summary %<>% mutate(fallecidos.prop = fallecidos/fallecidos.total.provincia)

 # Data 2 plot
 data2plot <- covid19.ar.summary %>% filter(residencia_provincia_nombre %in% covid19.ar.provincia.summary.100.confirmed$residencia_provincia_nombre)

 #Plot of deaths share
 covidplot <-
    data2plot %>%
    ggplot(aes(x = edad.rango, y = fallecidos.prop, fill = edad.rango)) +
    geom_bar(stat = "identity") + facet_wrap(~residencia_provincia_nombre, ncol = 2, scales = "free_y") +
    labs(title = "Proporción de muertos por rango etario\n en provincias > 100 confirmados")
 covidplot <- setupTheme(covidplot, report.date = report.date, x.values = NULL, x.type = NULL,
                      total.colors = length(unique(data2plot$edad.rango)),
                      data.provider.abv = "@msalnacion", base.size = 6)
 # Proporción de muertos por rango etario
 covidplot


 # UCI rate
 covidplot <- data2plot %>%
   ggplot(aes(x = edad.rango, y = cuidado.intensivo.porc, fill = edad.rango)) +
   geom_bar(stat = "identity") + facet_wrap(~residencia_provincia_nombre, ncol = 2, scales = "free_y") +
    labs(title = "Porcentaje de pacientes en Unidades de Cuidados Intensivos por rango etario\n en provincias > 100 confirmados")
 covidplot <- setupTheme(covidplot, report.date = report.date, x.values = NULL, x.type = NULL,
                      total.colors = length(unique(data2plot$edad.rango)),
                      data.provider.abv = "@msalnacion", base.size = 6)
 covidplot

 # ventilator rate
 covidplot <- data2plot %>%
   ggplot(aes(x = edad.rango, y = respirador.porc, fill = edad.rango)) +
   geom_bar(stat = "identity") +
   facet_wrap(~residencia_provincia_nombre, ncol = 2, scales = "free_y") +
   labs(title = "Porcentaje de pacientes con requerimientos de respirador mecánico por rango etario\n en provincias > 100 confirmados")
 covidplot <- setupTheme(covidplot, report.date = report.date, x.values = NULL, x.type = NULL,
                      total.colors = length(unique(data2plot$edad.rango)),
                      data.provider.abv = "@msalnacion", base.size = 6)
 covidplot

 # fatality rate

 covidplot <- data2plot %>%
  ggplot(aes(x = edad.rango, y = letalidad.min.porc, fill = edad.rango)) +
  geom_bar(stat = "identity") +
  facet_wrap(~residencia_provincia_nombre, ncol = 2, scales = "free_y") +
  labs(title = "Porcentaje de letalidad por rango etario\n en provincias > 100 confirmados")
 covidplot <- setupTheme(covidplot, report.date = report.date, x.values = NULL, x.type = NULL,
                      total.colors = length(unique(data2plot$edad.rango)),
                      data.provider.abv = "@msalnacion", base.size = 6)
 covidplot

```



# Generar diferentes agregaciones y guardar csv / Generate different aggregations

```R
output.dir <- "~/.R/COVID19AR/"
dir.create(output.dir, showWarnings = FALSE, recursive = TRUE)
exportAggregatedTables(covid19.curator, output.dir = output.dir,
                       aggrupation.criteria = list(provincia_residencia = c("residencia_provincia_nombre"),
                                                   provincia_localidad_residencia = c("residencia_provincia_nombre", "residencia_departamento_nombre"),
                                                   provincia_residencia_sexo = c("residencia_provincia_nombre", "sexo"),
                                                   edad_rango_sexo = c("edad.rango", "sexo"),
                                                   provincia_residencia_edad_rango = c("residencia_provincia_nombre", "edad.rango"),
                                                   provincia_residencia_sepi_apertura = c("residencia_provincia_nombre", "sepi_apertura"),
                                                   provincia_residencia = c("residencia_provincia_nombre", "residencia_departamento_nombre", "sepi_apertura"),
                                                   provincia_residencia_fecha_apertura = c("residencia_provincia_nombre", "fecha_apertura")))
                                                   
                                                  
```
All this tables are accesible at [COVID19ARdata](https://github.com/rOpenStats/COVID19ARdata/tree/master/curated) 

# How to Cite This Work


Citation
```
Alejandro Baranek, COVID19AR, 2020. URL: https://github.com/rOpenStats/COVID19AR
```

```bibtex
BibTex
@techreport{baranek2020Covid19AR,
Author = {Alejandro Baranek},
Institution = {rOpenStats},
Title = {COVID19AR: a package for analysing Argentina COVID-19 outbreak},
Url = {https://github.com/rOpenStats/COVID19AR},
Year = {2020}}
```
